---
title: "CR_Portfolio_2"
author: "Seunghun Lee"
date: "Oct 12, 2020"
output: 
  pdf_document:
    toc: true
    toc_depth: 2
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include=FALSE}
library(dplyr)
library(tidyr)
library(ggplot2)
library(magrittr)
library(psych)
```

## Research Question
Researchers were interested in whether the correlation between first mathematics exam score (G1) and final mathematics exam  score (G3) was stronger for male students than female students in a secondary education of Portuguese school. They obtained a random sample of 395 students (208 females, 187 males) from an independent dataset and investigated their research question using an alpha of 0.05 (P. Cortez and A. Silva, 2008).

```{r}
# https://archive.ics.uci.edu/ml/datasets/student+performance
dat <- read.table('./data/student-mat.csv',sep=";",header=TRUE)
# Subset necessary attributes
dat_f <- dat %>%
  select(sex, G1, G3) %>%
  filter(sex == 'F')
summary(dat_f)

dat_m <- dat %>%
  select(sex, G1, G3) %>%
  filter(sex  == 'M')
summary(dat_m)

dat %>%
  select(sex, G1, G3) %>%
  ggplot(aes(G1, G3, color = sex)) +
  geom_point() +
  geom_smooth(method = 'lm', se = FALSE) +
  facet_grid(cols = vars(sex)) +
  labs(title = 'First Exam Score (G1) vs Final Exam Score (G3)') +
  theme(legend.position = "none")
```

## Hypotheses
$H_{1}$: The correlation between first exam score (G1) and final exam score (G3) was stronger for male than female
($\rho_{M} - \rho_{F} > 0$) \
$H_{0}$: The correlation between first exam score (G1) and final score (G3) was not stronger for male than female
($\rho_{M} - \rho_{F} \le 0$) \

## Critical test statistic
```{r}
round(abs(qnorm(0.05)), 3)
```
$\alpha$ = 0.05, one-tailed, Critical Z-statistics = 1.645

## Test statistic
1. Build a function for independent correlation test
```{r, indent = '   '}
indep_cor_test <- function(data, group, var1, var2){
  #Get group levels
  group_levels <- data %>%
    pull({{group}}) %>%
    levels()
  #Correlation in group 1
  data1 <- data %>%
    select({{group}}, {{var1}}, {{var2}}) %>%
    filter({{group}} == group_levels[1]) 
  cor1 <- data1 %>%
    mutate(var1 = var1 <- {{var1}},
           var2 = var2 <- {{var2}}) %$%
    cor(var1, var2)
  #Correlation in group 2
  data2 <- data %>%
    select({{group}}, {{var1}}, {{var2}}) %>%
    filter({{group}} == group_levels[2])
  cor2 <- data2 %>%
    mutate(var1 = var1 <- {{var1}},
           var2 = var2 <- {{var2}}) %$%
    cor(var1, var2)
  #Independent correlation test
  rtest <- r.test(n = nrow(data1), n2 = nrow(data2), cor1, cor2, twotailed = FALSE)
  #print the results
  print(paste("The correlation in", group_levels[1]))
  print(cor1)
  print(paste("The correlation in", group_levels[2]))
  print(cor2)
  rtest
  }
```

* Computation by the function above
```{r, indent = '   '}
indep_cor_test(dat, sex, G1, G3)
```
* Result sample statistic Z value 1.62 < Critical Z-statistic 1.645\

2. Computation by hand
```{r, indent = '   '}
manual_z <- function(n1, cor1, n2, cor2){
  z1 <- 0.5*log((1+cor1)/(1-cor1))
  z2 <- 0.5*log((1+cor2)/(1-cor2))
  
  z <- ((z2 - z1) - 0)/sqrt((1/(n1-3))+(1/(n2-3)))
  print(abs(z))
}

manual_z(208, 0.7721856, 187, 0.8307415)
```
* Result sample statistic Z value 1.623 < Critical Z-statistic 1.645\


## Conclusion
* Fail to reject the null hypothesis\
* The correlation between G1 and G3 was not significantly stronger for male than for female\
* [$r_{F}$ = 0.77, $n_{F}$ = 208, $r_{M}$ = 0.83, $n_{M}$ = 187, Z = 1.62, p > 0.05]
