---
title: "Portfolio of Research Questions"
author: "Seunghun Lee"
date: "Sep 22, 2020"
output: 
  pdf_document:
    toc: true
    toc_depth: 2
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Research Question
Researchers were interested in whether the correlation between temperature of each day and the number of registered shared-bike users were positivley related. They obtained a random sample of 731 days and their values for 
temperature (Celcius), which was derived via (t-t_min)/(t_max-t_min), t_min = -8, t_max = +39 (only in hourly scale), and for registered users and investigated their research question using an alpha of 0.05. (Fanaee-T, Hadi and Gama, Joao, 2013)

```{r}
# https://archive.ics.uci.edu/ml/datasets/bike+sharing+dataset
day <- read.csv('./data/day.csv') 
day_temp_reg <- subset(day, select = c('temp', 'registered')) # subsetting the necessary columns
head(day_temp_reg)
summary(day_temp_reg)
```

## Hypotheses
H1: There is a positive relationship betwen temperature and the number of registered bike user.
$\rho > 0$ \
H0: There is not a positive relationship between temperature and the number of registered bike user.
$\rho \le 0$ \

## Critical test statistic
df = 731 - 2 = 729, $\alpha$-level = 0.05, one-tailed,
```{r}
abs(qt(0.05, 729))
```
Critical t(0.05, df = 729) = 1.647

## Test statistic
```{r}
# Defining necessary variables
temp <- day$temp 
registered <- day$registered
```

1. Computation by hand
```{r, indent = '   '}
tr_byhand <- function(data1, data2){
  r <- cor(data1, data2)
  Sr <- sqrt((1-r^2)/(length(data1)-2))
  Tr <- r/Sr
  print(c(r = r, t = Tr))
}
tr_byhand(temp, registered)
# r = 0.54, t = 17.323 
```

2. Test statistic via `cor.test` fuction
```{r, indent = '   '}
cor.test(temp, registered, alternative = 'greater')
```

## Conclusion
* Reject the null hypothesis (H0)\
* There is a significant positive relationship between temperature and the number of registered users \
* [r= 0.54, t(729) = 17.323, p < 0.05]
